Review of Edge Detectors

### 2015-02-13 edge-detection-review-3rd-revision-v3.tex
final version

\myn{\label{first:roberts}}The Roberts edge detector was introduced in 1963 by L.G. Roberts in the context of 3D reconstrution \cite{im_proc:segmentation:roberts:1963:boundary_extraction}, and it was based on the computation of the first derivative of the image. The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
Digital images are commonly seen as a discretization of a continuous image, where a pixel is evenly sampled on a grid. If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$. When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
Thus, in the discrete space the first derivative of the image is computed as:
\begin{equation}
\label{eq:roberts1}
	f_x = \frac{\partial f(x,y)}{\partial x} \cong f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	f_y = \frac{\partial f(x,y)}{\partial y} \cong f(i,j+1) - f(i,j).
\end{equation}

@@

The algorithm is divided in two steps, each one described later:
\begin{enumerate}
	\item Grayscale conversion of the input image (see section \ref{sec:appendix1}).
	\item Convolution of the image with:
	\begin{itemize}
		\item a Laplacian of Gaussian (LoG) kernel, (or)
		\item a Gaussian kernel and then a Laplacian operator.
	\end{itemize}
	\item Search of zero crossing points in the filtered image.
\end{enumerate}

@@

\myn{\label{second:haralick:fitting:statement}} The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
\begin{align}
	\label{eq:bicubic:long}
	f(x,y) = k_1 + k_2 (x-x_0) + k_3(y-y_0) + k_4(x-x_0)^2 + k_5(x-x_0)(y-y_0)  +\\
	 k_6(y-y_0) ^2 + k_7(x-x_0)^3 + k_8(x-x_0)^2(y-y_0)  + k_9(x-x_0)(y-y_0) ^2 + k_{10}(y-y_0) ^3 \\
\end{align}
To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other suitable fitting technique.\\

@@

\myn{\label{second:haralick:analytical:coordinates}} 
\begin{align*}
	x = \rho\sin{\theta} \\
	y = \rho\cos{\theta} 
\end{align*}

@@

\myn{\label{second:haralick:analytical:first}} If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain

@@

Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
(first odd integer greater than $6\sigma=24$).
\myn{\label{math:common:gauss}} 
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{kernel_gaussian.pdf}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. It is easy to see that the selected value of $n$ is 
large enough to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{figure}

@@

Now the LoG kernel is generated by sampling the function defined in equation \ref{eq:log_function}. 
Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the values $\sigma=4$ and $n=31$.
\myn{\label{math:common:log}} 
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{kernel_log.pdf}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of $n$ is sufficient 
to have a good approximation of the LoG function in the kernel, but it is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{figure}

@@

Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. It is clear that we must use a larger kernel size in the LoG function case (approximately 18\% larger). For 
example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
first odd integer greater than $6\sigma$, which is $25$, and for the case of LoG function, would 
be $29$.
% Especificar cual es el criterio para el tamaño del LoG.
% mirar en el software si no es mejor quitar la dependencia con n, y que se decida internamente el tamaño
% del kernel en funcion de sigma.
\myn{\label{math:common:comparison}} 
\begin{figure}[ht]
	\centering
	\includegraphics[width=1.0\textwidth]{kernels.pdf}
	\caption{Comparison of the Gaussian and LoG functions.}
	\label{fig:kernels}
\end{figure}

@@

\subsection{Effect of the scale parameter}
\myn{\label{results:parameter:scale}}In this section we will discuss the effect of the parameters for each of the reviewed algorithms. As a general rule, they all take one threshold parameter which acts as a scale parameter. As Figure \ref{fig:results:parameters:roberts:parameter} shows for the \emph{Roberts} algorithm, the bigger that threshold is, the coarser is the edge map obtained.

@@

\subsection{Effect of the smoothing parameter}
\myn{\label{results:parameter:smooth}}As we stated in previous section, all the reviewed algorithms share a scale parameter (threshold), however, the Marr-Hildreth Algorithm also has a smoothing step which adds an additional parameter $\sigma$, which controls the amount of regularization. In Figure \ref{fig:results:parameters:marr:sigma} we show the results using different  $\sigma$ values. As sigma increases more spurious edges are removed and the boundaries of the objects look smoother. However, if we keep increasing it we start removing meaningful edges (see the cloth band on the hat of Lena).

>>>>

The Roberts edge detector was introduced in 1963 by L.G. Roberts in the context of 3D reconstrution \cite{im_proc:segmentation:roberts:1963:boundary_extraction}, and it was based on the computation of the first derivative of the image. The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
Digital images are commonly seen as a discretization of a continuous image, where a pixel is evenly sampled on a grid. If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$. When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
Thus, in the discrete space the first derivative of the image is computed as:
\begin{equation}
\label{eq:roberts1}
	f_x = \frac{\partial f(x,y)}{\partial x} \cong f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	f_y = \frac{\partial f(x,y)}{\partial y} \cong f(i,j+1) - f(i,j).
\end{equation}

@@

The algorithm is divided in two steps, each one described later:
\begin{enumerate}
	\item Grayscale conversion of the input image.
	\item Convolution of the image with:
	\begin{itemize}
		\item a Laplacian of Gaussian (LoG) kernel, (or)
		\item a Gaussian kernel and then a Laplacian operator.
	\end{itemize}
	\item Search of zero crossing points in the filtered image.
\end{enumerate}

@@

The surrounding neighborhood of a point $(x_0,y_0)$ in the image $f$ is approximated using the following bi-cubic polynomial
\begin{align}
	\label{eq:bicubic:long}
	f(x,y) = k_1 + k_2 (x-x_0) + k_3(y-y_0) + k_4(x-x_0)^2 + k_5(x-x_0)(y-y_0)  +\\
	 k_6(y-y_0) ^2 + k_7(x-x_0)^3 + k_8(x-x_0)^2(y-y_0)  + k_9(x-x_0)(y-y_0) ^2 + k_{10}(y-y_0) ^3 \\
\end{align}
To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other suitable fitting technique.\\

@@

\begin{align*}
	x = \rho\sin{\theta} \\
	y = \rho\cos{\theta} 
\end{align*}

@@

If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain

@@

Figure \ref{fig:gaussian_kernel} shows a Gaussian kernel, generated with $\sigma = 4$ and $n = 25$ 
(first odd integer greater than $6\sigma=24$).
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{kernel_gaussian.pdf}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. It is easy to see that the selected value of $n$ is 
large enough to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{figure}

@@

Now the LoG kernel is generated by sampling the function defined in equation \ref{eq:log_function}. 
Figure \ref{fig:log_kernel} shows a LoG kernel, generated using the values $\sigma=4$ and $n=31$.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{kernel_log.pdf}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of $n$ is sufficient 
to have a good approximation of the LoG function in the kernel, but it is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{figure}

@@

Figure \ref{fig:kernels} shows both functions generated with the same value of $\sigma$. It is clear that we must use a larger kernel size in the LoG function case (approximately 18\% larger). For 
example, using $\sigma=4$, the optimum value for $n$ in the case of the Gaussian function is the 
first odd integer greater than $6\sigma$, which is $25$, and for the case of LoG function, would 
be $29$.
% Especificar cual es el criterio para el tamaño del LoG.
% mirar en el software si no es mejor quitar la dependencia con n, y que se decida internamente el tamaño
% del kernel en funcion de sigma.

\begin{figure}[ht]
	\centering
	\includegraphics[width=1.0\textwidth]{kernels.pdf}
	\caption{Comparison of the Gaussian and LoG functions.}
	\label{fig:kernels}
\end{figure}

@@

\subsection{Effect of the scale parameter}
In this section we will discuss the effect of the parameters for each of the reviewed algorithms. As a general rule, they all take one threshold parameter which acts as a scale parameter. As Figure \ref{fig:results:parameters:roberts:parameter} shows for the \emph{Roberts} algorithm, the bigger that threshold is, the coarser is the edge map obtained.

@@

\subsection{Effect of the smoothing parameter}
As we stated in previous section, all the reviewed algorithms share a scale parameter (threshold), however, the Marr-Hildreth Algorithm also has a smoothing step which adds an additional parameter $\sigma$, which controls the amount of regularization. In Figure \ref{fig:results:parameters:marr:sigma} we show the results using different  $\sigma$ values. As sigma increases more spurious edges are removed and the boundaries of the objects look smoother. However, if we keep increasing it we start removing meaningful edges (see the cloth band on the hat of Lena).
